---
title: "ST 540 hw4"
author: "Robin Baldeo"
date: "February 3, 2019"
output: word_document
---


```{r message = F, error=FALSE, warning=F,echo = T}
library(purrr)
library(tidyr)
library(dplyr)
library(knitr)
```

```{r}

#raw data for study
player<- c("Russell Westbrook","James Harden", "Kawhi Leonard", "Lebron James", "Isaiah Thomas", "Stephen Curry", "Giannis Antetokounmpo", "John Wall", "Anthony Davis", "Kevin Durant")
overall<- c(.845, .847, .880, .674, .909, .898, .770, .801, .802, .875)
cMakes<- c(64, 72, 55, 27, 75, 24, 28, 66, 40, 13)
cAtt<- c(75,95,63, 39, 83, 26, 41, 82, 54, 16)
id<- c(1:10)

#construct list of all vectors
data<- list( "player" = player, "overall"= overall, "cMakes" = cMakes, "cAtt"= cAtt, "id" = id)
```


##(A)
My prior is a Beta(1,1) because we are supposed to use a uniform prior with a binominal likelihood. Since this is a conjugate prior the result of this is a Beta(y + 1, n-y + 1) posterior. 

##(B)
```{r}
#theta
m<- seq(0,1, by = .01)

#function for posterior
betaPost<-function( x,n){
  a= x +1
  b = n-x + 1
  cbind("x" = m, "y" = dbeta(m, a ,b ))%>%return()
}
 
#create a list of all the posteriror to be plotted
clutch<- pmap(data, ~betaPost( ..3, ..4))

#plotting posteriror
op <- par(pty="m", mfrow=c(5, 2), mar=c(4.2, 4.2, 1, 1))

  for(i in 1:length(player)){
      plot(x = clutch[[i]][,1],y = clutch[[i]][,2], type = "l",col = "red", xlab = "Theta", ylab= "Post", main = player[i])
  }

par(op)
```

##(C)
```{r}

#function for summary
postSum<-function( x,n, nam, over){
  a= x +1
  b = n-x + 1
  #mean
  pMean=a/(a+b)
  #std
  pStd = sqrt((a*b)/(((a+b)**2)*(a+b+1)))
  #credible interval
  ci = qbeta(c(.025, .975), a,b)
  #lower bound credible interval
  lowCi=ci[1]
  #upper bound credible interval
  hiCi=ci[2]
  #portion less than .05
  less= pbeta(.5,a,b)
  cbind("name" = nam, "overAll"= over, "pMean" = pMean, "pStd" = pStd, "lowCi"= lowCi, "hiCi"= hiCi, "less"= less)%>%
    as.data.frame%>%
      return()
}

#creating dataframe with summary 
df_c<- pmap_df(data, ~postSum(..3,..4, ..5, ..2))

#adding players names 
df_c$name<- player[df_c$name]

df_c%>%kable(caption = "Summary Of Player's Posterior")
```


##(D)
```{r, warning=F}

#checking to see if the overall matches the posterior
# df_d<- pmap_df(data, ~{cbind("name" = ..5, "pBelow" = pbeta(..2,..3+ 1,..4 - ..3 + 1), "pAbove" = 1- pbeta(..2,..3+ 1,..4 - ..3 + 1))}%>%as.data.frame)

df_d<- pmap_df(data, ~{cbind("name" = ..5, "pBelow" = pbeta(..2,..3,..4 ), "pAbove" = 1- pbeta(..2,..3,..4 - ..3))}%>%as.data.frame)


#adding names and orginal overall to table
df_d$overAll<- overall[df_d$name]
df_d$name<- player[df_d$name]


df_d[-3]%>%
  arrange(pBelow)%>% 
    kable(caption = "%Percentage above and below Overall")


```

There is a high probability the overall matches the clutch for James Harden. With Kawhi Leonard, Isaiah Thomas, Kevin Durant, Anthony Davis, and Giannis Antetokounmpo the `pBelow` gives probability that the overall scores are less than the clutch percentage. While for Lebron, Stephen and Russel their probability suggest that clutch percentage should be greater than the overall.

##(E)
I don't think the data is sensitive to this prior since we are told that we are dealing with a large sample. So given the large sample size our posterior mean gets closer to the sample proportion. Therefore I think the prior has minimal effect. 



##MCMC

```{r}
#seed so experiment can be reproduced
set.seed(100)

#function that allows for random draws
mcmc<- function(x,n){
  a= x +1
  b = n-x + 1
  rbeta(100000, a,b)%>% return()
}

# Using the data data frame with raw information, random draws were made for each of the 10 players
df_e<- pmap(data, ~mcmc(..3,..4))

#a matrix to hold all the comparison scores for each player
l<-matrix(nrow= 10, ncol = 10)
#comparison 
for(i in 1:10){
  var<- double(10)
  for(j in 1:10){
    #all players were compared to each other and 
    #scores were recorded
    l[i,j]= (mean(df_e[[i]]> df_e[[j]]))
  }
}

#transofm each player results so it can be analysed row wise
tr<- t(l)

#max score of each player was determined and the player was recorded
rank<- as.integer(rep(0,10))
per<- as.double(rep(0, 10))
for(i in 1:10){
  rank[i]= tr[i,]%>%which.max()
  per[i] = tr[i,]%>%max()
  
}

#summary of best player against each player
results<- data.frame("Player" = player, "Best Score" = player[rank], "Percentage" = per)
results%>%kable(caption = "Best Player for one on one comparison")

```

Using simulation we see that Isaiah Thomas has the best free throw when compared to all players except Stephen Curry. Stephen Curry is the only player to beat Isaiah Thomas, therefore Stephen Curry is the best player. 


